Parallel Global Optimization for Non-Convex Mixed-Integer Problems

### 2019-03-27 2019/RSD_Integer/RSD2019.tex

Let us consider a multiextremal optimization problem in the form
\begin{gather}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},\\
D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.\nonumber
\end{gather}
Suppose, that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
the left-hand sides $g_j(y), \; 1\leq j \leq m$, of the constraints satisfy the Lipschitz condition
\[ \left|g_j(y')-g_j(y'')\right| \leq L_j \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq j \leq m+1, \]
with constants $L_j, \; 1 \leq j \leq m+1$, respectively,  
and may be multiextremal.

Using a continuous single-valued mapping $y(x)$  (Peano-type space-filling curve, or \textit{evolvent}) of the interval $[0,1]$ onto $D$, a multidimensional problem (\ref{problem}) can be reduced to a one-dimensional problem
\begin{equation}\label{problem1}
\varphi(y(x^\ast))= \min \left\{g_{m+1}(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\end{equation}
The reduction of dimensionality matches the multidimensional problem with a Lipschitzian objective function and constraints with a one-dimensional problem where the respective functions satisfy the uniform H\"older condition (see \cite{Strongin2000}).

An efficient global search algorithm (GSA) for solving the constrained optimization problem 
(\ref{problem1}) has been developed at University of Nizhni Novgorod.

%Описание индексного алгоритма

A detailed description 
of the algorithm and the corresponding convergence theory are presented in 
\cite{Strongin2000}.

\section{Global search algorithm for mixed-integer problems}

@@

Рассмотрим теперь соответствие 
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
\[
g_j(x) = g_j(Y(x),u_{E(x)+1}), x\in[0,S],
\]
имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
Используя введенные обозначения можно переформулировать исходную задачу как
\[
\min \left\{g_{m+1}(x): x \in [0,S], \; g_j(x) \leq 0, \; 1 \leq j \leq m\right\}.
\]

@@

Applying the global search algorithm to solving the problem (\ref{problem_is1}), we will find 
the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted in 
the subproblem, the solving of which corresponds to the solving of the initial 
problem (\ref{problem_i}). In the rest subproblems, a only minor part of trials will be performed 
since the solutions of these subproblems are the locally optimal ones.
%with respect to the solution of the $s$-th subproblem. 
All the above is confirmed by Fig. \ref{fig:1}, where the  
points of trials executed in the course of solving this problem are denoted by the dashes.

Thus, we have constructed the \textit{Mixed-Integer Global Search Algorithm} (MIGSA) based on the  
reduction of the mixed-integer non-convex optimization problem to the non-convex optimization problem. 
%It is possible to prove that the convergence conditions for this algorithm will follow from the convergence conditions of its prototype (GSA).

@@

Let us compare the proposed MIGSA method with a genetic algorithm for solving the mixed-integer optimization problems implemented in Matlab Global Optimization Toolbox \cite{Matlab}. In Table 
\ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
problems by these methods are presented. For both methods, the same accuracy of search $10^{-2}$ were used. All numerical experiments were conducted on a 
computer with Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM under MS Windows 10. The 
results of experiments have demonstrated the advantage to the MIGSA method in the number 
of iterations as well as in the execution time.

\begin{table}
	\caption{Comparison of MIGSA and GA}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c}{ GA } & & \multicolumn{2}{c}{MIGSA} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & $k$ & $t$ & & $k$ & $t$  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		 Problem 2 \cite{Floudas}&	481 &	0.0601 & &	417 &	0.04 \\
%		 Problem 3 \cite{Floudas}& 	1821 &	0.1130 & & 3324 &	0.107 \\
		 Problem 6 \cite{Floudas}&	641 &	0.0510 & &	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	481 &	0.1378 & &	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	481 &	0.0473 & &	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	841 &	0.0736 & & 372	 &	0.017 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

>>>>

A constrained global optimization problem can be formulated as follows
%Let us consider the $N$-dimensional global optimization problem
\begin{gather}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.\label{D}
\end{gather}
The objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand 
sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
\[
\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
\]
with a priori unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is 
assumed that the functions $g_i(y)$ are defined and computable only at the points $y \in D$ 
satisfying the conditions
\begin{equation}\label{g_k}
g_k(y) \leq 0, \; 1 \leq k < i.
\end{equation}

By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval 
$[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the 
minimum in (\ref{problem}) by solving the one-dimensional problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]
Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are 
considered in \cite{Strongin2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined 
and computable in the subranges 
\[
Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

These conditions allows us to introduce a classification of the points $x \in [0,1]$ according to 
the number $\nu (x)$ of the constraints computed at this point. The \textit{index} $\nu(x)$ can 
also be defined by the conditions
\begin{equation}\label{nu}
g_i(y(x)) \leq 0, \; 1 \leq i < \nu, \; g_\nu(y(x))>0,
\end{equation}
where the last inequality is inessential if $\nu=m+1$.

In the dimensionality reduction scheme considered here, a multidimensional problem with 
Lipschitzian functions is juxtaposed with a one-dimensional problem, where the corresponding 
functions satisfy uniform H{\"o}lder condition (see \cite{Strongin2013}), i.e.,
\[
\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 
1\leq i \leq m+1.
\]
Here, $N$ is the dimensionality of the initial multidimensional problem and the coefficients 
$H_i$ are related to the Lipschitz constant $L_i$ of the initial problem as $H_i \leq 2L_i 
\sqrt{N+3}$.

Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm 
will consist of the following sequence of operations:
\begin{itemize}
	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping 
$y(x)$;
	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from 
(\ref{nu}). 
\end{itemize}
The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial 
outcome}.

An efficient parallel index algorithm (PIA) for solving the constrained optimization problem 
(\ref{problem}) has been developed at University of Nizhni Novgorod.

Suppose we have $p \geq 1$  computational elements (e.g., accelerator cores), which can be used to run $p$ trials simultaneously. In the first iteration of the method, $p$ trials are run in parallel at various random points $x^i\in(0,1)$, $1\leq i \leq p$. 
Suppose $n \geq 1$  iterations of the method have been completed, and as a result of which, trials were carried out in $k=k(n)$ points $x^i, 1\leq i \leq k$. Then the points $x^{k+1},...,x^{k+p}$  of the search trials in the next $(n+1)$-th iteration will be determined according to the rules below.

\begin{enumerate}
\item 
Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest coordinate values, i.e.
\begin{equation}\label{Eq:17}
0=x_0<x_1<...<x_i<...<x_k<x_{k+1}=1,
\end{equation}
and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{Eq:16}) , calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
\item
Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of problem constraints fulfilled at these points, by building the sets
\begin{equation}\label{Eq:18}
I_\nu = \left\{i: 1 \leq i \leq k,\ \nu = \nu(x_i)\right\},\ 1 \leq \nu \leq m+1,
\end{equation}
containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are matched to an additional set $I_0={0,k+1}$.

Identify the maximum current value of the index
\begin{equation}\label{Eq:19}
M=\max \left\{\nu = \nu(x_i), \ 1\leq i \leq k\right\}.
\end{equation}
\item
For all values of $\nu, \ 1\leq \nu \leq m+1$, calculate the values  
\begin{equation}\label{Eq:20}
\mu_\nu = \max \left\{ \frac{\left|z_i-z_j\right|}{\left(x_i-x_j\right)^{1/N}} : i,j \in I_\nu, j<i\right\}.
\end{equation}
If the set $I_\nu$ contains less than two elements or $\mu_\nu$ from (\ref{Eq:20}) equals zero, then assume $\mu_\nu=1$.
\item
For all non-empty sets $I_\nu$, $1 \leq \nu \leq m+1$, determine the values
\begin{equation}\label{Eq:21}
  z^\ast_\nu =  
   \begin{cases}
    -\epsilon_\nu,  \nu < M, \\
    \min{\left\{g_\nu(x_i):i\in I_\nu\right\}}, \nu = M,
   \end{cases}
\end{equation}
where $M$ is the maximum current value of the index, and the vector
\begin{equation}\label{Eq:22}
\epsilon _R=\left(\epsilon_1,...,\epsilon_m\right),
\end{equation}
with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the algorithm.
\item
For each interval $(x_{i-1},x_i)$,$1 \leq i \leq k+1$, calculate the \textit{characteristic} $R(i)$: 
\[
R(i)=\Delta_i+ \frac{(z_i-z_{i-1})^2}{(r_\nu\mu_\nu)^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu=\nu(x_{i-1})=\nu(x_i),
\]
\[
R(i)= 2\Delta_i-4\frac{z_i-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu(x_{i-1})<\nu(x_i)=\nu,
\]
\[
R(i)= 2\Delta_i-4\frac{z_{i-1}-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu = \nu(x_{i-1})>\nu(x_i).
\]
where $\Delta_i=(x_i-x_{i-1})^{1/N}$, and the values $r_\nu>1, 1\leq\nu\leq m+1$, are used as parameters in the algorithm.
\item
Reorder the characteristics $R(i)$, $1\leq i \leq k+1$, from highest to lowest 	
\begin{equation}\label{Eq:23}
R(t_1)\geq R(t_2)\geq ... \geq R(t_{k})\geq R(t_{k+1})
\end{equation}
and choose $p$ largest characteristics with interval numbers $t_j, 1\leq j \leq p$.
\item
Carry out $p$ new trials in parallel at the points $x^{k+j}, 1 \leq j \leq p$, calculated by the formulae
\begin{eqnarray*}
& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-1})=\nu(x_{t_j})=\nu. \\
\end{eqnarray*}

%Добавить результаты испытаний в информационную базу алгоритма, и перейти к правилу 1.
\end{enumerate}

The algorithm stops if the condition $\Delta_{t_j}\leq \epsilon$ becomes true for at least one number $t_j, 1\leq j \leq p$; here  $\epsilon>0$ has an order of magnitude of the desired coordinate accuracy.

A detailed description of the algorithm convergence theory is presented in \cite{Strongin2000}.

\section{Parallel algorithm for mixed-integer problems}

@@

Рассмотрим теперь соответствие 
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
\[
g_j(x) = g_j(Y(x),u_{E(x)+1}), x\in[0,S],
\]
имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
%значения функций и индексов в данных точках - не определены!
Используя введенные обозначения можно переформулировать исходную задачу как
\begin{equation}\label{problem_is1}
\min \left\{g_{m+1}(x): x \in [0,S], \; g_j(x) \leq 0, \; 1 \leq j \leq m\right\}.
\end{equation}

@@

Applying the parallel index algorithm to solving the problem (\ref{problem_is1}), we will find 
the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted in 
the subproblem, the solving of which corresponds to the solving of the initial 
problem (\ref{problem_i}). In the rest subproblems, a only minor part of trials will be performed 
since the solutions of these subproblems are the locally optimal ones.
%with respect to the solution of the $s$-th subproblem. 
All the above is confirmed by Fig. \ref{fig:1}, where the  
points of trials executed in the course of solving this problem are denoted by the dashes.

Thus, we have constructed the \textit{Mixed-Integer Parallel Index Algorithm} (MIPIA) based on the  
reduction of the mixed-integer non-convex optimization problem to the non-convex optimization problem. 
%It is possible to prove that the convergence conditions for this algorithm will follow from the convergence conditions of its prototype (GSA).

@@

Let us compare proposed MIPIA with a genetic algorithm for solving the mixed-integer global optimization problems implemented in Matlab Global Optimization Toolbox \cite{Matlab}. In Table 
\ref{tab:1}, the numbers of trials required for solving the known test mixed-integer 
problems by these methods are presented. For both methods, the same accuracy of search $10^{-2}$ were used. All numerical experiments were conducted on a 
computer with Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM under MS Windows 10. The 
results of experiments have demonstrated the advantage of MIPIA in the number 
of iterations as well as in the execution time.

\begin{table}
	\caption{Comparison of MIPIA and GA}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c}{ GA } & & \multicolumn{2}{c}{MIGSA} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & $k$ & $t$ & & $k$ & $t$  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		 Problem 2 \cite{Floudas}&	481 &	0.0601 & &	417 &	0.04 \\
%		 Problem 3 \cite{Floudas}& 	1821 &	0.1130 & & 3324 &	0.107 \\
		 Problem 6 \cite{Floudas}&	641 &	0.0510 & &	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	481 &	0.1378 & &	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	481 &	0.0473 & &	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	841 &	0.0736 & & 372	 &	0.017 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}
